---
title: "Quick & dirty sneak peak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    df_print: paged
    highlight: tango
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r knitr, include = FALSE}
knitr::opts_chunk$set(
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    message        = FALSE,
    warning        = FALSE,
    fig.width      = 10,
    fig.height     = 8
)
```

```{r logs}
log <- file(snakemake@log[[1]], open = "wt")
sink(log)
sink(log, type = "message")
```

```{r libraries}
library(Seurat)
library(tidyverse)
library(viridis)
source("workflow/scripts/functions.R")
```

```{r input-files-and-params}
input_file   <- snakemake@input[[1]]
marker_genes <- snakemake@params[["marker_genes"]]
species      <- snakemake@params[["species"]]
cluster_degs <- snakemake@output[["cluster_degs"]]
sample_degs  <- snakemake@output[["sample_degs"]]
```

```{r load-seurat}
seurat <- readRDS(input_file)
```

```{r seurat-pipeline, results = 'hide'}
# CC genes
data(cc.genes.updated.2019)
if (species == "mouse") {
  s.genes   <- str_to_title(cc.genes.updated.2019$s.genes)
  g2m.genes <- str_to_title(cc.genes.updated.2019$g2m.genes)
} else {
  s.genes   <- cc.genes.updated.2019$s.genes
  g2m.genes <- cc.genes.updated.2019$g2m.genes
}

# Pipeline params
n_pcs <- 25
k     <- 20

# Seurat pipeline
seurat <- seurat %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(npcs = n_pcs) %>%
  FindNeighbors(dims = 1:n_pcs, k.param = k) %>%
  FindClusters() %>%
  RunUMAP(dims = 1:n_pcs, n.neighbors = k, min.dist = 0.3) %>%
  CellCycleScoring(s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```

# Mitochondrial and Ribosomal RNA by sample

Mitochondrial RNA by sample:

```{r, fig.width = 7.5, fig.height = 4}
VlnPlot(seurat, features = c("percent.mt"), pt.size = 0) + theme(legend.position = "none")
```

Ribosomal RNA by sample:

```{r, fig.width = 7.5, fig.height = 4}
VlnPlot(seurat, features = c("percent.ribo"), pt.size = 0) + theme(legend.position = "none")
```


# Mitochondrial, Ribosomal RNA by cluster

Mitochondrial RNA by cluster:

```{r, fig.width = 7.5, fig.height = 4}
VlnPlot(seurat, features = c("percent.mt"), pt.size = 0, group.by = "seurat_clusters") + theme(legend.position = "none")
```

Ribosomal RNA by cluster:

```{r, fig.width = 7.5, fig.height = 4}
VlnPlot(seurat, features = c("percent.ribo"), pt.size = 0, group.by = "seurat_clusters") + theme(legend.position = "none")
```


# UMAP plots {.tabset .tabset-fade}

```{r umap-plots, results='asis', fig.width=7, fig.height=5.5}
seurat_sample_plt  <- DimPlot(seurat, group.by = "sample")
seurat_cluster_plt <- DimPlot(seurat, group.by = "seurat_clusters")
seurat_cc_plt      <- DimPlot(seurat, group.by = "Phase")
seurat_RNA_plt     <- FeaturePlot(seurat, features = "nFeature_RNA", max.cutoff = "q99") + scale_color_viridis()
seurat_Count_plt   <- FeaturePlot(seurat, features = "nCount_RNA", max.cutoff = "q99") + scale_color_viridis()
seurat_mt_plt      <- FeaturePlot(seurat, features = "percent.mt", max.cutoff = "q99") + scale_color_viridis()
seurat_ribo_plt    <- FeaturePlot(seurat, features = "percent.ribo", max.cutoff = "q99") + scale_color_viridis()

l <- list(
  UMAP_sample    = seurat_sample_plt,
  UMAP_cluster   = seurat_cluster_plt,
  UMAP_CellCycle = seurat_cc_plt,
  UMAP_nGenes    = seurat_RNA_plt,
  UMAP_nCounts   = seurat_Count_plt,
  UMAP_mt        = seurat_mt_plt,
  UMAP_ribo      = seurat_ribo_plt
)

in_tabs(l, level = 1L)
```

# Expression of specific genes {.tabset .tabset-fade}

```{r marker-genes, results='asis', fig.width = 7.5, fig.height = 5.5}
detected_genes <- GetAssayData(object = seurat, slot = "data") %>% rownames()

plots <- map(
  marker_genes,
  \(gene) plot_detected_genes(seurat, detected_genes, gene)
)

in_tabs(plots, labels = marker_genes, level = 1L)
```


# Differential expression between clusters {.tabset .tabset-fade}

```{r diff-exp-clust, results='asis'}
cluster_markers <- FindAllMarkers(seurat, only.pos = T, logfc.threshold = 0.25)
write_tsv(cluster_markers, cluster_degs)

top_9_markers <- cluster_markers %>% 
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice(1:9) %>%
  split(f = as.factor(paste0("Cluster ", .$cluster))) %>%
  map(pull, gene)

plots <- map(
  top_9_markers,
  \(genes) FeaturePlot(seurat, features = genes, ncol = 3) & scale_color_viridis()
)

in_tabs(plots, level = 1L)
```

# Differential expression between samples {.tabset .tabset-fade}

```{r diff-exp-sample, results='asis', fig.width = 9.5, fig.height = 13}
Idents(seurat)  <- seurat$sample
sample_markers <- FindAllMarkers(seurat, only.pos = T, logfc.threshold = 0.25)
write_tsv(cluster_markers, sample_degs)
Idents(seurat)  <- seurat$seurat_clusters

top_9_markers <- sample_markers %>% 
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice(1:9) %>%
  split(f = as.factor(.$cluster)) %>%
  map(pull, gene)

plots <- map(
  top_9_markers,
  \(genes) FeaturePlot(seurat, features = genes, ncol = 3) & scale_color_viridis()
)

in_tabs(plots, level = 1L)
```